# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r echo=TRUE}
rawdata<-read.csv("activity.csv", na.strings="NA")
```


## What is mean total number of steps taken per day?

```{r echo=TRUE}
stepsbyday<-aggregate(rawdata$steps, by=list(date = rawdata$date), FUN="sum")
plot(strptime(stepsbyday$date, "%Y-%m-%d"), stepsbyday$x, type="hist", xlab="Date", ylab = "Number of Steps")
title("Total number of steps by day")
meanbyday <- mean(stepsbyday$x, na.rm=TRUE)
medianbyday<-median(stepsbyday$x, na.rm = TRUE)

```

Mean steps count per day: `r meanbyday`  
Median steps count per day: `r medianbyday`

## What is the average daily activity pattern?


```{r echo=TRUE}
fdata<-subset(rawdata, !is.na(steps), select=c(steps, date, interval))
stepsbyinterval<-aggregate(fdata$steps, by=list(interval = fdata$interval), FUN="mean")
plot(stepsbyinterval$interval, stepsbyinterval$x, type="l", xlab="interval", ylab = "Number of Steps")
title("Average number of steps by 5-minutes interval")
maxinteravl<-stepsbyinterval$interval[max(stepsbyinterval$x)==stepsbyinterval$x]
```

Interval with maximum number of steps: `r maxinteravl`

```{r echo=TRUE}
nadata<-subset(rawdata, is.na(steps), select=c(steps, date, interval))
misval <- dim(nadata)[1]
maxinteravl<-stepsbyinterval$interval[max(stepsbyinterval$x)==stepsbyinterval$x]
```


## Imputing missing values



```{r echo=TRUE}
nadata<-subset(rawdata, is.na(steps), select=c(date, interval))
misval <- dim(nadata)[1]
```

Number of missing values is `r misval`

Now we replcaing missing values by mean of 5-minute intervals calculated withowt missing data

```{r echo=TRUE}
tmp<-merge(stepsbyinterval, nadata, by="interval")
updata<-data.frame(steps=tmp$x, date = tmp$date, interval = tmp$interval)
newdata<-rbind(updata, fdata)
newstepsbyday<-aggregate(newdata$steps, by=list(date = newdata$date), FUN="sum")
plot(strptime(newstepsbyday$date, "%Y-%m-%d"), newstepsbyday$x, type="hist", xlab="Date", ylab = "Number of Steps")
title("Total number of steps by day with recovered missing data")
newmeanbyday <- mean(newstepsbyday$x)
newmedianbyday<-median(newstepsbyday$x)
```

Mean steps count per day after recover missing values: `r newmeanbyday` (difference is `r newmeanbyday-meanbyday`)  
Median steps count per day after recover missing values: `r newmedianbyday` (difference is `r newmedianbyday-medianbyday`)


## Are there differences in activity patterns between weekdays and weekends?

```{r echo=TRUE}
wd0<-as.POSIXlt(strptime(newdata$date, "%Y-%m-%d"))$wday == 0
wd6<-as.POSIXlt(strptime(newdata$date, "%Y-%m-%d"))$wday == 6
fwd<-as.factor(wd0|wd6)
levels(fwd) = c("weekday", "weekend")
wdata<-data.frame(steps=newdata$steps, date=newdata$date, interval=newdata$interval,weekend=fwd)
stepsbyintervalwd<-aggregate(wdata$steps, by=list(interval = wdata$interval, weekend = wdata$weekend), FUN="mean")
library(lattice)
xyplot(stepsbyintervalwd$x~stepsbyintervalwd$interval|stepsbyintervalwd$weekend, type='l', xlab = "interval", layout = c(1,2))

